Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-07-04 WSSSPE2_report/WSSSPE2_review.tex
Section 7: minor changes.

Despite agreement that this body of disciplinary software needed improvement,
there was general agreement that such improvement needed to be understood in
terms of the scientific productivity that could be achieved. A software maturity
model is needed for science software, but it needs to be introduced in a way
that fits the culture of science, which largely thinks of software as a tool,
rather than a product itself. The group was in general consensus that more
widespread training in software practice is needed within the domain sciences,
and several of the participants were involved in efforts along these lines.
Participants felt that projects that build a sense of community via training in
software and technical practices would have the most success in changing
practices in that community, but that there was a need for a managed
introduction of these practices. Participants also recognized that these could
not be one-time, one-off training opportunities, as software and technologies
change rapidly over time. For example, while today Software Carpentry is focused
on teaching version control via Git, there has been a rapid evolution from RCS
to CVS to SVN to Git over a short time frame, and thus communities should expect
to need to train for adaptability and a changing technology tool chain, rather
than assume these technologies stay fixed. Thus, although short term training
that introduces specific tools was considered highly valuable, these trainings
were also not considered sufficient to engender the changes in software practice
that were deemed necessary. Combining the need for changes in practice to be
introduced incrementally with the need to to minimize the divergence of training
from direct science goals and the difficulties of training for a rapidly
evolving technology space, participants concluded that multiple training efforts
that targeted different parts of the spectrum were needed. Short-term courses
introducing immediately useful skills needs to be offered alongside more
in-depth courses on software engineering and practice that allow students to
adapt to a changing landscape.

>>>>

%Despite agreement that this body of disciplinary software needed improvement,
There was general agreement that this body of disciplinary software improvement needed to be understood in
terms of the scientific productivity that could be achieved. A software maturity
model is needed for science software, but it needs to be introduced in a way
that fits the culture of science, which largely thinks of software as a tool,
rather than a product itself. The group was in general consensus that more
widespread training in software practice is needed within the domain sciences,
and several of the participants were involved in efforts along these lines.
Participants felt that projects that build a sense of community via training in
software and technical practices would have the most success in changing
practices in that community, but that there was a need for a managed
introduction of these practices. Participants also recognized that these could
not be one-time, one-off training opportunities, as software and technologies
change rapidly over time. For example, while today Software Carpentry is focused
on teaching version control via Git, there has been a rapid evolution from RCS
to CVS to SVN to Git over a short time frame, and thus communities should expect
the need to train for adaptability and a changing technology tool chain, rather
than assume that these technologies will stay fixed. Thus, although short term training
that introduces specific tools was considered highly valuable, these trainings
were also not considered sufficient to engender the changes in software practice
that were deemed necessary. Combining the need for changes in practice to be
introduced incrementally with the need to minimize the divergence of training
from direct science goals and the difficulties of training for a rapidly
evolving technology space, participants concluded that multiple training efforts
that targeted different parts of the spectrum were needed. Short-term courses
introducing immediately useful skills needs to be offered alongside more
in-depth courses on software engineering and practice that allow students to
adapt to a changing landscape.
